---
title: "STAT 331 Portfolio"
author: "Laura Miller"
format: html 
embed-resources: true
layout: margin-left
editor: visual
execute: 
  eval: false
  echo: true
---

[**My Grade:**]{.underline} I believe my grade equivalent to course work evidenced below to be an \_\_.

[**Learning Objective Evidence:**]{.underline} In the code chunks below, provide code from Lab or Challenge assignments where you believe you have demonstrated proficiency with the specified learning target. Be sure to specify **where** the code came from (e.g., Lab 4 Question 2).

## Working with Data

**WD-1: I can import data from a *variety* of formats (e.g., csv, xlsx, txt, etc.).**

-   `csv` Example 1: Lab 2 Question 1

```{r}
#| label: wd-1-csv-1
surveys <- read_csv(file = here("Week 2", "surveys.csv"))
```

-   `csv` Example 2: Lab 3 Question 2

```{r}
#| label: wd-1-csv-2
teacher_evals <- read_csv(file = here("Week 3", "teacher_evals.csv"))
```

-   `xlsx`

```{r}
#| label: wd-1-xlsx

```

**WD-2: I can select necessary columns from a dataset.**

-   Example selecting specified columns: Lab 3 Question 5

```{r}
#| label: wd-2-ex-1
teacher_evals_clean <- teacher_evals |>
  rename(sex = gender) |>
  filter(no_participants >= 10) |>
  mutate(teacher_id = as.character(teacher_id),
         course_id = as.character(course_id)) |>
  select(course_id, teacher_id, question_no, no_participants, resp_share, 
         SET_score_avg, percent_failed_cur, academic_degree, seniority, sex)
```

-   Example removing specified columns

```{r}
#| label: wd-2-ex-2

```

-   Example selecting columns based on logical values (e.g., `starts_with()`, `ends_with()`, `contains()`, `where()`): Lab 3 Question 1

```{r}
#| label: wd-2-ex-3
teacher_evals_compare <- teacher_evals |>
  filter(question_no == 903) |>
  mutate(SET_level = if_else(SET_score_avg >= 4, "excellent", "standard"),
         sen_level = case_when(seniority <= 4 ~ "junior",
                               seniority > 4 ~ "senior")) |>
  select(course_id, SET_level, sen_level)
```

**WD-3: I can filter rows from a dataframe for a *variety* of data types (e.g., numeric, integer, character, factor, date).**

-   Numeric Example 1: Challenge 3 Question 1

```{r}
#| label: wd-3-numeric-ex-1
teacher_evals_compare <- teacher_evals |>
  filter(question_no == 903) |>
  mutate(SET_level = if_else(SET_score_avg >= 4, "excellent", "standard"),
         sen_level = case_when(seniority <= 4 ~ "junior",
                               seniority > 4 ~ "senior")) |>
  select(course_id, SET_level, sen_level)
```

-   Numeric Example 2: Lab 4 Question 6

```{r}
#| label: wd-3-numeric-ex-1
infant_cost_2018 <- ca_childcare |>
  filter(study_year == 2018) |>
  group_by(census_region) |>
  summarize(med_mc_infant = median(mc_infant, na.rm = TRUE),
            .groups = "drop") |>
  arrange(desc(med_mc_infant))
```

-   Character Example 1 (any context): Lab 4 Question 2

```{r}
#| label: wd-3-character
ca_childcare <- childcare_costs |>
  left_join(x = childcare_costs,
            y = counties,
            by = join_by(county_fips_code == county_fips_code)) |>
  filter(state_abbreviation == "CA") 
```

-   Character Example 2 (example must use functions from **stringr**): Lab 3 Question 12

```{r}
#| label: wd-3-string
resp_avg <- teacher_evals_clean |>
  filter(academic_degree %in% c("dr", "prof")) |>
  group_by(teacher_id, seniority, sex) |>
  summarise(n_courses = n_distinct(course_id),
            avg_resp = mean(resp_share, na.rm = TRUE),
            .groups = "drop")
```

-   Date (example must use functions from **lubridate**): Lab 5

```{r}
#| label: wd-3-date
suspect_check_in <- get_fit_now_check_in |>
  inner_join(get_fit_now_member, by = c("membership_id" = "id")) |>
  mutate(membership_id = as.character(membership_id),
         check_in_date = ymd(as.character(check_in_date))) |>
  filter(check_in_date == as.Date("2018-01-09"), 
         str_detect(membership_id, "^48Z"))
suspect_check_in
```

**WD-4: I can modify existing variables and create new variables in a dataframe for a *variety* of data types (e.g., numeric, integer, character, factor, date).**

-   Numeric Example 1: Lab 5

```{r}
#| label: wd-4-numeric-ex-1
suspect_check_in <- get_fit_now_check_in |>
  inner_join(get_fit_now_member, by = c("membership_id" = "id")) |>
  mutate(membership_id = as.character(membership_id),
         check_in_date = ymd(as.character(check_in_date))) |>
  filter(check_in_date == as.Date("2018-01-09"), 
         str_detect(membership_id, "^48Z"))
suspect_check_in
```

-   Numeric Example 2: Lab 3 Question 1

```{r}
#| label: wd-4-numeric-ex-2
teacher_evals_compare <- teacher_evals |>
  filter(question_no == 903) |>
  mutate(SET_level = if_else(SET_score_avg >= 4, "excellent", "standard"),
         sen_level = case_when(seniority <= 4 ~ "junior",
                               seniority > 4 ~ "senior")) |>
  select(course_id, SET_level, sen_level)
```

-   Factor Example 1 (renaming levels): Lab 4 Question 7

```{r}
#| label: wd-4-factor-ex-1
ca_childcare_long <- ca_childcare |>
  select(census_region, county_name, study_year, mc_infant, mc_toddler, mc_preschool) |>
  mutate(study_year = as.numeric(study_year)) |>
  pivot_longer(cols = c(mc_infant, mc_toddler, mc_preschool),
               names_to = "age_group",
               values_to = "median_weekly_cost") |>
  mutate(age_group = recode(age_group,
                            mc_infant = "Infant",
                            mc_toddler = "Toddler",
                            mc_preschool = "Preschool"),
         age_group = factor(age_group, levels = c("Infant", "Toddler", "Preschool"))) |>
  group_by(age_group, census_region) |>
  mutate(census_region = fct_reorder(census_region, median_weekly_cost, .fun = max)) |>
  ungroup()
```

-   Factor Example 2 (reordering levels): Lab 4 Question 7

```{r}
#| label: wd-4-factor-ex-2
ca_childcare_long <- ca_childcare |>
  select(census_region, county_name, study_year, mc_infant, mc_toddler, mc_preschool) |>
  mutate(study_year = as.numeric(study_year)) |>
  pivot_longer(cols = c(mc_infant, mc_toddler, mc_preschool),
               names_to = "age_group",
               values_to = "median_weekly_cost") |>
  mutate(age_group = recode(age_group,
                            mc_infant = "Infant",
                            mc_toddler = "Toddler",
                            mc_preschool = "Preschool"),
         age_group = factor(age_group, levels = c("Infant", "Toddler", "Preschool"))) |>
  group_by(age_group, census_region) |>
  mutate(census_region = fct_reorder(census_region, median_weekly_cost, .fun = max)) |>
  ungroup()
```

-   Character (example must use functions from **stringr**): Lab 4 Question 4

```{r}
#| label: wd-4-string
ca_childcare <- ca_childcare |> 
  mutate(county_name = str_remove(county_name, " County"),
    census_region = fct_collapse(county_name, "Superior California" = superior_counties, 
                                  "North Coast" = north_coast_counties,
                                  "San Francisco Bay Area" = san_fran_counties,
                                  "Northern San Joaquin Valley" = n_san_joaquin_counties,
                                  "Central Coast" = central_coast_counties,
                                  "Southern San Joaquin Valley" = s_san_joaquin_counties,
                                  "Inland Empire" = inland_counties,
                                  "Los Angeles" = la_county,
                                  "Orange County" = orange_county,
                                  "San Diego - Imperial" = san_diego_imperial_counties))
```

-   Date (example must use functions from **lubridate**): Lab 5

```{r}
#| label: wd-4-date
suspect_check_in <- get_fit_now_check_in |>
  inner_join(get_fit_now_member, by = c("membership_id" = "id")) |>
  mutate(membership_id = as.character(membership_id),
         check_in_date = ymd(as.character(check_in_date))) |>
  filter(check_in_date == as.Date("2018-01-09"), 
         str_detect(membership_id, "^48Z"))
suspect_check_in
```

**WD-5: I can use mutating joins to combine multiple dataframes.**

-   `left_join()` Example 1: Lab 4 Question 2

```{r}
#| label: wd-5-left-ex-1
ca_childcare <- childcare_costs |>
  left_join(x = childcare_costs,
            y = counties,
            by = join_by(county_fips_code == county_fips_code)) |>
  filter(state_abbreviation == "CA") 

ca_childcare
```

-   `right_join()` Example 1

```{r}
#| label: wd-5-right

```

-   `left_join()` **or** `right_join()` Example 2: Lab 4 Question 3

```{r}
#| label: wd-5-left-right-ex-2
ca_childcare <- ca_childcare |>
  left_join(x = ca_childcare,
            y = tax_rev,
            by = join_by(county_name == entity_name,
                         study_year == year))

ca_childcare
```

-   `inner_join()` Example 1: Lab 5

```{r}
#| label: wd-5-inner-ex-1
suspect_check_in <- get_fit_now_check_in |>
  inner_join(get_fit_now_member, by = c("membership_id" = "id")) |>
  mutate(membership_id = as.character(membership_id),
         check_in_date = ymd(as.character(check_in_date))) |>
  filter(check_in_date == as.Date("2018-01-09"), 
         str_detect(membership_id, "^48Z"))
suspect_check_in
```

-   `inner_join()` Example 2: Lab 5

```{r}
#| label: wd-5-inner-ex-2
suspect <- get_fit_now_check_in |>
  inner_join(get_fit_now_member, by = c("membership_id" = "id")) |>
  mutate(check_in_date = ymd(check_in_date)) |>
  filter(check_in_date == as.Date("2018-01-09"),
         str_detect(membership_id, "^48Z")) |>
  inner_join(person, by = c("person_id" = "id"))

suspect <- suspect |>
  inner_join(drivers_license, by = c("license_id" = "id")) |>
  filter(str_detect(plate_number, "H42W"))
suspect
```

**WD-6: I can use filtering joins to filter rows from a dataframe.**

-   `semi_join()`

```{r}
#| label: wd-6-semi

```

-   `anti_join()`

```{r}
#| label: wd-6-anti

```

**WD-7: I can pivot dataframes from long to wide and visa versa**

-   `pivot_longer()` : Challenge 4 Summary Statistics

```{r}
#| label: wd-7-long
med_childcare_cost_by_setting <- ca_childcare |>
  select(study_year, mcsa, mfccsa) |>
  pivot_longer(cols = c(mcsa, mfccsa),
               names_to = "setting",
               values_to = "cost") |>
  group_by(setting, study_year) |>
  summarise(median_cost = median(cost, na.rm = TRUE),
            .groups = "drop") |>
  pivot_wider(names_from = study_year,
              values_from = median_cost,
              names_prefix = "year_") |>
  arrange(setting, desc(year_2018))

med_childcare_cost_by_setting
```

-   `pivot_wider()` : Challenge 4 Summary Statistics

```{r}
#| label: wd-7-wide
med_childcare_cost_by_setting <- ca_childcare |>
  select(study_year, mcsa, mfccsa) |>
  pivot_longer(cols = c(mcsa, mfccsa),
               names_to = "setting",
               values_to = "cost") |>
  group_by(setting, study_year) |>
  summarise(median_cost = median(cost, na.rm = TRUE),
            .groups = "drop") |>
  pivot_wider(names_from = study_year,
              values_from = median_cost,
              names_prefix = "year_") |>
  arrange(setting, desc(year_2018))

med_childcare_cost_by_setting
```

## Reproducibility

**R-1: I can create professional looking, reproducible analyses using RStudio projects, Quarto documents, and the here package.**

The following assignments satisfy the above criteria:

-   Lab 1
-   Lab 2
-   Lab 3
-   Lab 4
-   Challenge 4

**R-2: I can write well documented and tidy code.**

-   Example of **ggplot2** plotting: Lab 1 Question 6

```{r}
#| label: r-2-1
library(tidyverse)

ggplot(data = ToothGrowth, 
       mapping = aes(x = supp, y = len)) +
  geom_boxplot() +
  labs(x = "Supplement", y = "Length of Teeth (mm)")
ToothGrowth %>% 
  mutate(dose = as.factor(dose)) %>% 
  ggplot(mapping = aes(x = dose, y = len)) +
  geom_boxplot() +
  labs(x = "Dose of Supplement (mg/day)", y = "Length of Teeth (mm)")
```

-   Example of **dplyr** pipeline: Lab 3 Question 1

```{r}
#| label: r-2-2
teacher_evals_compare <- teacher_evals |>
  filter(question_no == 903) |>
  mutate(SET_level = if_else(SET_score_avg >= 4, "excellent", "standard"),
         sen_level = case_when(seniority <= 4 ~ "junior",
                               seniority > 4 ~ "senior")) |>
  select(course_id, SET_level, sen_level)

teacher_evals_compare
```

-   Example of function formatting

```{r}
#| label: r-2-3

```

**R-3: I can write robust programs that are resistant to changes in inputs.**

-   Example (any context)

```{r}
#| label: r-3-example

```

-   Example (function stops)

```{r}
#| label: r-3-function-stops

```

## Data Visualization & Summarization

**DVS-1: I can create visualizations for a *variety* of variable types (e.g., numeric, character, factor, date)**

-   At least two numeric variables

```{r}
#| label: dvs-1-num

```

-   At least one numeric variable and one categorical variable

```{r}
#| label: dvs-2-num-cat

```

-   At least two categorical variables

```{r}
#| label: dvs-2-cat

```

-   Dates (time series plot)

```{r}
#| label: dvs-2-date

```

**DVS-2: I use plot modifications to make my visualization clear to the reader.**

-   I can modify my plot theme to be more readable

```{r}
#| label: dvs-2-ex-1

```

-   I can modify my colors to be accessible to anyone's eyes

```{r}
#| label: dvs-2-ex-2

```

-   I can modify my plot titles to clearly communicate the data context

```{r}
#| label: dvs-2-ex-3

```

-   I can modify the text in my plot to be more readable

```{r}
#| label: dvs-2-ex-4

```

-   I can reorder my legend to align with the colors in my plot

```{r}
#| label: dvs-2-ex-5

```

**DVS-3: I show creativity in my visualizations**

-   I can use non-standard colors (Example 1)

```{r}
#| label: dvs-3-1-ex-1

```

-   I can use non-standard colors (Example 2)

```{r}
#| label: dvs-3-1-ex-2

```

-   I can use annotations (e.g., `geom_text()`)

```{r}
#| label: dvs-3-2

```

-   I can choose creative geometries (e.g., `geom_segment()`, `geom_ribbon)()`)

```{r}
#| label: dvs-3-3

```

**DVS-4: I can calculate numerical summaries of variables.**

-   Example using `summarize()`

```{r}
#| label: dvs-4-summarize

```

-   Example using `across()`

```{r}
#| label: dvs-4-across

```

**DVS-5: I can find summaries of variables across multiple groups.**

-   Example 1

```{r}
#| label: dvs-5-1

```

-   Example 2

```{r}
#| label: dvs-5-2

```

**DVS-6: I can create tables which make my summaries clear to the reader.**

-   I can modify my column names to clearly communicate the data context

```{r}
#| label: dvs-6-ex-1

```

-   I can modify the text in my table to be more readable (e.g., bold face for column headers)

```{r}
#| label: dvs-6-ex-2

```

-   I can arrange my table to have an intuitive ordering

```{r}
#| label: dvs-6-ex-3

```

**DVS-7: I show creativity in my tables.**

-   I can use non-default colors

```{r}
#| label: dvs-7-ex-1

```

-   I can modify the layout of my table to be more readable (e.g., `pivot_longer()` or `pivot_wider()`)

```{r}
#| label: dvs-7-ex-2

```

## Program Efficiency

**PE-1: I can write concise code which does not repeat itself.**

-   using a single function call with multiple inputs (rather than multiple function calls)

```{r}
#| label: pe-1-one-call

```

-   using `across()`

```{r}
#| label: pe-1-across

```

-   using functions from the `map()` family

```{r}
#| label: pe-1-map-1

```

**PE-2: I can write functions to reduce repetition in my code.**

-   Example 1: Function that operates on vectors

```{r}
#| label: pe-2-1

```

-   Example 2: Function that operates on data frames

```{r}
#| label: pe-2-2

```

-   Example 3: Function that operates on vectors *or* data frames

```{r}
#| label: pe-2-3

```

**PE-3:I can use iteration to reduce repetition in my code.**

-   using `across()`

```{r}
#| label: pe-3-across

```

-   using a `map()` function with **one** input (e.g., `map()`, `map_chr()`, `map_dbl()`, etc.)

```{r}
#| label: pe-3-map-1

```

-   using a `map()` function with **more than one** input (e.g., `map_2()` or `pmap()`)

```{r}
#| label: pe-3-map-2

```

**PE-4: I can use modern tools when carrying out my analysis.**

-   I can use functions which are not superseded or deprecated

```{r}
#| label: pe-4-1

```

-   I can connect a data wrangling pipeline into a `ggplot()`

```{r}
#| label: pe-4-2

```

## Data Simulation & Statisical Models

**DSSM-1: I can simulate data from a *variety* of probability models.**

-   Example 1

```{r}
#| label: dsm-1-1

```

-   Example 2

```{r}
#| label: dsm-1-2

```

**DSSM-2: I can conduct common statistical analyses in R.**

-   Example 1

```{r}
#| label: dsm-2-1

```

-   Example 2

```{r}
#| label: dsm-2-2

```

-   Example 3

```{r}
#| label: dsm-2-3

```

## Revising My Thinking

<!-- How did you revise your thinking throughout the course? How did you revise your thinking on the code examples you have provided in your portfolio? -->

<!-- For the revisions included in your Portfolio, to help me understand the nature of your revisions, please denote somehow the feedback I provided you (e.g., boldface, italics, colored text) before your revisions. -->

## Extending My Thinking

<!-- How did you extended your thinking throughout the course? How did you extend your thinking on the code examples you have provided in your portfolio? -->

## Peer Support & Collaboration

<!-- Include an image or a description of feedback you gave that you are proud of (either in a peer review or in Discord). -->

<!-- Include a description of how you grew as a collaborator through the weekly pair programming activities.   -->
