---
title: "STAT 331 Portfolio"
author: "Laura Miller"
format: html 
embed-resources: true
layout: margin-left
editor: visual
execute: 
  eval: false
  echo: true
---

[**My Grade:**]{.underline} I believe my grade equivalent to course work evidenced below to be an \_B-\_.

[**Learning Objective Evidence:**]{.underline} In the code chunks below, provide code from Lab or Challenge assignments where you believe you have demonstrated proficiency with the specified learning target. Be sure to specify **where** the code came from (e.g., Lab 4 Question 2).

## Working with Data

**WD-1: I can import data from a *variety* of formats (e.g., csv, xlsx, txt, etc.).**

-   `csv` Example 1: Lab 2 Question 1

```{r}
#| label: wd-1-csv-1
surveys <- read_csv(file = here("Week 2", "surveys.csv"))
```

-   `csv` Example 2: Lab 3 Question 2

```{r}
#| label: wd-1-csv-2
teacher_evals <- read_csv(file = here("Week 3", "teacher_evals.csv"))
```

-   `xlsx`

```{r}
#| label: wd-1-xlsx

```

**WD-2: I can select necessary columns from a dataset.**

-   Example selecting specified columns: Lab 3 Question 5

```{r}
#| label: wd-2-ex-1
teacher_evals_clean <- teacher_evals |>
  rename(sex = gender) |>
  filter(no_participants >= 10) |>
  mutate(teacher_id = as.character(teacher_id),
         course_id = as.character(course_id)) |>
  select(course_id, teacher_id, question_no, no_participants, resp_share, 
         SET_score_avg, percent_failed_cur, academic_degree, seniority, sex)
```

-   Example removing specified columns

```{r}
#| label: wd-2-ex-2

```

-   Example selecting columns based on logical values (e.g., `starts_with()`, `ends_with()`, `contains()`, `where()`): Lab 3 Question 1

```{r}
#| label: wd-2-ex-3
teacher_evals_compare <- teacher_evals |>
  filter(question_no == 903) |>
  mutate(SET_level = if_else(SET_score_avg >= 4, "excellent", "standard"),
         sen_level = case_when(seniority <= 4 ~ "junior",
                               seniority > 4 ~ "senior")) |>
  select(course_id, SET_level, sen_level)
```

**WD-3: I can filter rows from a dataframe for a *variety* of data types (e.g., numeric, integer, character, factor, date).**

-   Numeric Example 1: Challenge 3 Question 1

```{r}
#| label: wd-3-numeric-ex-1
teacher_evals_compare <- teacher_evals |>
  filter(question_no == 903) |>
  mutate(SET_level = if_else(SET_score_avg >= 4, "excellent", "standard"),
         sen_level = case_when(seniority <= 4 ~ "junior",
                               seniority > 4 ~ "senior")) |>
  select(course_id, SET_level, sen_level)
```

-   Numeric Example 2: Lab 4 Question 6

```{r}
#| label: wd-3-numeric-ex-1
infant_cost_2018 <- ca_childcare |>
  filter(study_year == 2018) |>
  group_by(census_region) |>
  summarize(med_mc_infant = median(mc_infant, na.rm = TRUE),
            .groups = "drop") |>
  arrange(desc(med_mc_infant))
```

-   Character Example 1 (any context): Lab 4 Question 2

```{r}
#| label: wd-3-character
ca_childcare <- childcare_costs |>
  left_join(x = childcare_costs,
            y = counties,
            by = join_by(county_fips_code == county_fips_code)) |>
  filter(state_abbreviation == "CA") 
```

-   Character Example 2 (example must use functions from **stringr**): Lab 3 Question 12

```{r}
#| label: wd-3-string
resp_avg <- teacher_evals_clean |>
  filter(academic_degree %in% c("dr", "prof")) |>
  group_by(teacher_id, seniority, sex) |>
  summarise(n_courses = n_distinct(course_id),
            avg_resp = mean(resp_share, na.rm = TRUE),
            .groups = "drop")
```

-   Date (example must use functions from **lubridate**): Lab 5

```{r}
#| label: wd-3-date
suspect_check_in <- get_fit_now_check_in |>
  inner_join(get_fit_now_member, by = c("membership_id" = "id")) |>
  mutate(membership_id = as.character(membership_id),
         check_in_date = ymd(as.character(check_in_date))) |>
  filter(check_in_date == as.Date("2018-01-09"), 
         str_detect(membership_id, "^48Z"))
suspect_check_in
```

**WD-4: I can modify existing variables and create new variables in a dataframe for a *variety* of data types (e.g., numeric, integer, character, factor, date).**

-   Numeric Example 1: Lab 5

```{r}
#| label: wd-4-numeric-ex-1
suspect_check_in <- get_fit_now_check_in |>
  inner_join(get_fit_now_member, by = c("membership_id" = "id")) |>
  mutate(membership_id = as.character(membership_id),
         check_in_date = ymd(as.character(check_in_date))) |>
  filter(check_in_date == as.Date("2018-01-09"), 
         str_detect(membership_id, "^48Z"))
suspect_check_in
```

-   Numeric Example 2: Lab 3 Question 1

```{r}
#| label: wd-4-numeric-ex-2
teacher_evals_compare <- teacher_evals |>
  filter(question_no == 903) |>
  mutate(SET_level = if_else(SET_score_avg >= 4, "excellent", "standard"),
         sen_level = case_when(seniority <= 4 ~ "junior",
                               seniority > 4 ~ "senior")) |>
  select(course_id, SET_level, sen_level)
```

-   Factor Example 1 (renaming levels): Lab 4 Question 7

```{r}
#| label: wd-4-factor-ex-1
ca_childcare_long <- ca_childcare |>
  select(census_region, county_name, study_year, mc_infant, mc_toddler, mc_preschool) |>
  mutate(study_year = as.numeric(study_year)) |>
  pivot_longer(cols = c(mc_infant, mc_toddler, mc_preschool),
               names_to = "age_group",
               values_to = "median_weekly_cost") |>
  mutate(age_group = recode(age_group,
                            mc_infant = "Infant",
                            mc_toddler = "Toddler",
                            mc_preschool = "Preschool"),
         age_group = factor(age_group, levels = c("Infant", "Toddler", "Preschool"))) |>
  group_by(age_group, census_region) |>
  mutate(census_region = fct_reorder(census_region, median_weekly_cost, .fun = max)) |>
  ungroup()
```

-   Factor Example 2 (reordering levels): Lab 4 Question 7

```{r}
#| label: wd-4-factor-ex-2
ca_childcare_long <- ca_childcare |>
  select(census_region, county_name, study_year, mc_infant, mc_toddler, mc_preschool) |>
  mutate(study_year = as.numeric(study_year)) |>
  pivot_longer(cols = c(mc_infant, mc_toddler, mc_preschool),
               names_to = "age_group",
               values_to = "median_weekly_cost") |>
  mutate(age_group = recode(age_group,
                            mc_infant = "Infant",
                            mc_toddler = "Toddler",
                            mc_preschool = "Preschool"),
         age_group = factor(age_group, levels = c("Infant", "Toddler", "Preschool"))) |>
  group_by(age_group, census_region) |>
  mutate(census_region = fct_reorder(census_region, median_weekly_cost, .fun = max)) |>
  ungroup()
```

-   Character (example must use functions from **stringr**): Lab 4 Question 4

```{r}
#| label: wd-4-string
ca_childcare <- ca_childcare |> 
  mutate(county_name = str_remove(county_name, " County"),
    census_region = fct_collapse(county_name, "Superior California" = superior_counties, 
                                  "North Coast" = north_coast_counties,
                                  "San Francisco Bay Area" = san_fran_counties,
                                  "Northern San Joaquin Valley" = n_san_joaquin_counties,
                                  "Central Coast" = central_coast_counties,
                                  "Southern San Joaquin Valley" = s_san_joaquin_counties,
                                  "Inland Empire" = inland_counties,
                                  "Los Angeles" = la_county,
                                  "Orange County" = orange_county,
                                  "San Diego - Imperial" = san_diego_imperial_counties))
```

-   Date (example must use functions from **lubridate**): Lab 5

```{r}
#| label: wd-4-date
suspect_check_in <- get_fit_now_check_in |>
  inner_join(get_fit_now_member, by = c("membership_id" = "id")) |>
  mutate(membership_id = as.character(membership_id),
         check_in_date = ymd(as.character(check_in_date))) |>
  filter(check_in_date == as.Date("2018-01-09"), 
         str_detect(membership_id, "^48Z"))
suspect_check_in
```

**WD-5: I can use mutating joins to combine multiple dataframes.**

-   `left_join()` Example 1: Lab 4 Question 2

```{r}
#| label: wd-5-left-ex-1
ca_childcare <- childcare_costs |>
  left_join(x = childcare_costs,
            y = counties,
            by = join_by(county_fips_code == county_fips_code)) |>
  filter(state_abbreviation == "CA") 

ca_childcare
```

-   `right_join()` Example 1

```{r}
#| label: wd-5-right

```

-   `left_join()` **or** `right_join()` Example 2: Lab 4 Question 3

```{r}
#| label: wd-5-left-right-ex-2
ca_childcare <- ca_childcare |>
  left_join(x = ca_childcare,
            y = tax_rev,
            by = join_by(county_name == entity_name,
                         study_year == year))

ca_childcare
```

-   `inner_join()` Example 1: Lab 5

```{r}
#| label: wd-5-inner-ex-1
suspect_check_in <- get_fit_now_check_in |>
  inner_join(get_fit_now_member, by = c("membership_id" = "id")) |>
  mutate(membership_id = as.character(membership_id),
         check_in_date = ymd(as.character(check_in_date))) |>
  filter(check_in_date == as.Date("2018-01-09"), 
         str_detect(membership_id, "^48Z"))
suspect_check_in
```

-   `inner_join()` Example 2: Lab 5

```{r}
#| label: wd-5-inner-ex-2
suspect <- get_fit_now_check_in |>
  inner_join(get_fit_now_member, by = c("membership_id" = "id")) |>
  mutate(check_in_date = ymd(check_in_date)) |>
  filter(check_in_date == as.Date("2018-01-09"),
         str_detect(membership_id, "^48Z")) |>
  inner_join(person, by = c("person_id" = "id"))

suspect <- suspect |>
  inner_join(drivers_license, by = c("license_id" = "id")) |>
  filter(str_detect(plate_number, "H42W"))
suspect
```

**WD-6: I can use filtering joins to filter rows from a dataframe.**

-   `semi_join()`

```{r}
#| label: wd-6-semi

```

-   `anti_join()`

```{r}
#| label: wd-6-anti

```

**WD-7: I can pivot dataframes from long to wide and visa versa**

-   `pivot_longer()` : Challenge 4 Summary Statistics

```{r}
#| label: wd-7-long
med_childcare_cost_by_setting <- ca_childcare |>
  select(study_year, mcsa, mfccsa) |>
  pivot_longer(cols = c(mcsa, mfccsa),
               names_to = "setting",
               values_to = "cost") |>
  group_by(setting, study_year) |>
  summarise(median_cost = median(cost, na.rm = TRUE),
            .groups = "drop") |>
  pivot_wider(names_from = study_year,
              values_from = median_cost,
              names_prefix = "year_") |>
  arrange(setting, desc(year_2018))

med_childcare_cost_by_setting
```

-   `pivot_wider()` : Challenge 4 Summary Statistics

```{r}
#| label: wd-7-wide
med_childcare_cost_by_setting <- ca_childcare |>
  select(study_year, mcsa, mfccsa) |>
  pivot_longer(cols = c(mcsa, mfccsa),
               names_to = "setting",
               values_to = "cost") |>
  group_by(setting, study_year) |>
  summarise(median_cost = median(cost, na.rm = TRUE),
            .groups = "drop") |>
  pivot_wider(names_from = study_year,
              values_from = median_cost,
              names_prefix = "year_") |>
  arrange(setting, desc(year_2018))

med_childcare_cost_by_setting
```

## Reproducibility

**R-1: I can create professional looking, reproducible analyses using RStudio projects, Quarto documents, and the here package.**

The following assignments satisfy the above criteria:

-   Lab 1
-   Lab 2
-   Lab 3
-   Lab 4
-   Challenge 4

**R-2: I can write well documented and tidy code.**

-   Example of **ggplot2** plotting: Lab 1 Question 6

```{r}
#| label: r-2-1
library(tidyverse)

ggplot(data = ToothGrowth, 
       mapping = aes(x = supp, y = len)) +
  geom_boxplot() +
  labs(x = "Supplement", y = "Length of Teeth (mm)")
ToothGrowth %>% 
  mutate(dose = as.factor(dose)) %>% 
  ggplot(mapping = aes(x = dose, y = len)) +
  geom_boxplot() +
  labs(x = "Dose of Supplement (mg/day)", y = "Length of Teeth (mm)")
```

-   Example of **dplyr** pipeline: Lab 3 Question 1

```{r}
#| label: r-2-2
teacher_evals_compare <- teacher_evals |>
  filter(question_no == 903) |>
  mutate(SET_level = if_else(SET_score_avg >= 4, "excellent", "standard"),
         sen_level = case_when(seniority <= 4 ~ "junior",
                               seniority > 4 ~ "senior")) |>
  select(course_id, SET_level, sen_level)

teacher_evals_compare
```

-   Example of function formatting

```{r}
#| label: r-2-3

```

**R-3: I can write robust programs that are resistant to changes in inputs.**

-   Example (any context)

```{r}
#| label: r-3-example

```

-   Example (function stops)

```{r}
#| label: r-3-function-stops

```

## Data Visualization & Summarization

**DVS-1: I can create visualizations for a *variety* of variable types (e.g., numeric, character, factor, date)**

-   At least two numeric variables: Lab 4 Question 8

```{r}
#| label: dvs-1-num
ggplot(ca_childcare, aes(x = mhi_2018, y = mc_infant)) +
  geom_point(alpha = 0.5, color = "blue") +
  geom_smooth(method = "lm", se = TRUE, color = "red", size = 1) +
  labs(x = "Median Household Income 2018 ($)",
       y = "Median Weekly Cost for Infant Center-Based Childcare ($)",
       title = "Relationship Between Household Income and Infant Childcare Costs in CA",
       color = "Region") +
  theme_bw() +
  theme(axis.title.y = element_text(size = 8),
        plot.title = element_text(size = 10, face = "bold"))
```

-   At least one numeric variable and one categorical variable: Lab 2 Question 16

```{r}
#| label: dvs-2-num-cat
ggplot(data = surveys, mapping = aes(x = weight, y = species)) +
  geom_boxplot(outlier.shape = NA) +
  labs(x = "Weight (g)", y = "Species", title = "Distribution of Species by Weight") +
  geom_jitter(color = "steelblue", alpha = 0.3)
```

-   At least two categorical variables: Challenge 2

```{r}
#| label: dvs-2-cat
ggplot(data = surveys, mapping = aes(x = species, y = weight, color = sex)) +
  geom_jitter(color = "green", alpha = 0.02) +
  geom_boxplot(outlier.shape = NA) +
  labs(x = "Species", y = "Weight (g)", title = "Distribution of Weight by Species") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = c("pink", "blue"))
```

-   Dates (time series plot): Challenge 4 Plot

```{r}
#| label: dvs-2-date
med_childcare_cost_by_setting <- ca_childcare |>
  mutate(study_year = as.integer(study_year)) |>
  select(study_year, mcsa, mfccsa) |>
  pivot_longer(cols = c(mcsa, mfccsa),
               names_to = "setting",
               values_to = "cost") |>
  group_by(setting, study_year) |>
  summarise(median_cost = median(cost, na.rm = TRUE),
            .groups = "drop")

ggplot(med_childcare_cost_by_setting, aes(x = study_year,
                                          y = median_cost,
                                          color = setting)) +
    geom_line(size = 1) +
    geom_point(size = 2, alpha = 0.5) +
    scale_x_continuous(breaks = 
                         seq(min(med_childcare_cost_by_setting$study_year),
                        max(med_childcare_cost_by_setting$study_year), by = 1)) +
    labs(title = "Median Full-Time Weekly Childcare Costs in CA by Setting",
         x = "Study Year",
         y = "Median Weekly Cost ($)",
         color = "Setting") +
    theme_bw()
```

**DVS-2: I use plot modifications to make my visualization clear to the reader.**

-   I can modify my plot theme to be more readable: Lab 4 Question 8

```{r}
#| label: dvs-2-ex-1
ggplot(ca_childcare, aes(x = mhi_2018, y = mc_infant)) +
  geom_point(alpha = 0.5, color = "blue") +
  geom_smooth(method = "lm", se = TRUE, color = "red", size = 1) +
  labs(x = "Median Household Income 2018 ($)",
       y = "Median Weekly Cost for Infant Center-Based Childcare ($)",
       title = "Relationship Between Household Income and Infant Childcare Costs in CA",
       color = "Region") +
  theme_bw() +
  theme(axis.title.y = element_text(size = 8),
        plot.title = element_text(size = 10, face = "bold"))
```

-   I can modify my colors to be accessible to anyone's eyes: Challenge 3 Question 2

```{r}
#| label: dvs-2-ex-2
ggplot(teacher_evals_compare, aes(x = sen_level, fill = SET_level)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("excellent" = "purple", "standard" = "orange")) +
  labs(x = "Years of Experience",
       y = "",
       title = "Evaluation of Teachers' Use of Activities",
       fill = "Evaluation Rating")
```

-   I can modify my plot titles to clearly communicate the data context: Challenge 2

```{r}
#| label: dvs-2-ex-3
ggplot(data = surveys, mapping = aes(x = species, y = weight, color = sex)) +
  geom_jitter(color = "green", alpha = 0.02) +
  geom_boxplot(outlier.shape = NA) +
  labs(x = "Species", y = "Weight (g)", title = "Distribution of Weight by Species") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = c("pink", "blue"))
```

-   I can modify the text in my plot to be more readable: Lab 4 Question 7

```{r}
#| label: dvs-2-ex-4
ca_childcare_long <- ca_childcare |>
  select(census_region, county_name, study_year, mc_infant, mc_toddler, mc_preschool) |>
  mutate(study_year = as.numeric(study_year)) |>
  pivot_longer(cols = c(mc_infant, mc_toddler, mc_preschool),
               names_to = "age_group",
               values_to = "median_weekly_cost") |>
  mutate(age_group = recode(age_group,
                            mc_infant = "Infant",
                            mc_toddler = "Toddler",
                            mc_preschool = "Preschool"),
         age_group = factor(age_group, levels = c("Infant", "Toddler", "Preschool"))) |>
  group_by(age_group, census_region) |>
  mutate(census_region = fct_reorder(census_region, median_weekly_cost, .fun = max)) |>
  ungroup()
        

ggplot(ca_childcare_long, aes(x = study_year, 
                              y = median_weekly_cost,
                              color = census_region,
                              group = interaction(census_region, county_name))) +
  geom_point(alpha = 0.3) +
  geom_smooth(aes(group = census_region), 
              method = "loess", 
              se = FALSE, 
              linewidth = 1) +
  facet_wrap(~ age_group, ncol = 3, scales = "fixed") +
  labs(x = "Study Year",
       y = "",
       title = "Weekly Median Price for Center-Based Childcare ($)",
       color = "California Region") +
  theme_bw(base_size = 10) +
  scale_y_continuous(limits = c(100, 500)) +
  scale_x_continuous(limits = c(2008, 2018),
                     breaks = seq(2008, 2018, by = 2)) +
  theme(axis.text.x = element_text(size = 6),
        strip.text = element_text(size = 10),
        legend.text = element_text(size = 9),
        legend.title = element_text(size = 10),
        panel.spacing = unit(0.5, "cm"))
```

-   I can reorder my legend to align with the colors in my plot: Lab 4 Question 7

```{r}
#| label: dvs-2-ex-5
ca_childcare_long <- ca_childcare |>
  select(census_region, county_name, study_year, mc_infant, mc_toddler, mc_preschool) |>
  mutate(study_year = as.numeric(study_year)) |>
  pivot_longer(cols = c(mc_infant, mc_toddler, mc_preschool),
               names_to = "age_group",
               values_to = "median_weekly_cost") |>
  mutate(age_group = recode(age_group,
                            mc_infant = "Infant",
                            mc_toddler = "Toddler",
                            mc_preschool = "Preschool"),
         age_group = factor(age_group, levels = c("Infant", "Toddler", "Preschool"))) |>
  group_by(age_group, census_region) |>
  mutate(census_region = fct_reorder(census_region, median_weekly_cost, .fun = max)) |>
  ungroup()
        

ggplot(ca_childcare_long, aes(x = study_year, 
                              y = median_weekly_cost,
                              color = census_region,
                              group = interaction(census_region, county_name))) +
  geom_point(alpha = 0.3) +
  geom_smooth(aes(group = census_region), 
              method = "loess", 
              se = FALSE, 
              linewidth = 1) +
  facet_wrap(~ age_group, ncol = 3, scales = "fixed") +
  labs(x = "Study Year",
       y = "",
       title = "Weekly Median Price for Center-Based Childcare ($)",
       color = "California Region") +
  theme_bw(base_size = 10) +
  scale_y_continuous(limits = c(100, 500)) +
  scale_x_continuous(limits = c(2008, 2018),
                     breaks = seq(2008, 2018, by = 2)) +
  theme(axis.text.x = element_text(size = 6),
        strip.text = element_text(size = 10),
        legend.text = element_text(size = 9),
        legend.title = element_text(size = 10),
        panel.spacing = unit(0.5, "cm"))
```

**DVS-3: I show creativity in my visualizations**

-   I can use non-standard colors (Example 1): Challenge 2

```{r}
#| label: dvs-3-1-ex-1
ggplot(data = surveys, mapping = aes(x = species, y = weight, color = sex)) +
  geom_jitter(color = "green", alpha = 0.02) +
  geom_boxplot(outlier.shape = NA) +
  labs(x = "Species", y = "Weight (g)", title = "Distribution of Weight by Species") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = c("pink", "blue"))
```

-   I can use non-standard colors (Example 2): Challenge 2

```{r}
#| label: dvs-3-1-ex-2
ggplot(data = surveys, mapping = aes(x = weight, y = species)) +
  geom_jitter(color = "steelblue", alpha = 0.1) +
  geom_boxplot(outlier.shape = NA) +
  labs(x = "Weight (g)", y = "Species", title = "Weight Distribution by Species") +
  theme(panel.grid.minor = element_blank()) +
  theme(axis.title.y = element_text(margin = margin(r = 10)))
```

-   I can use annotations (e.g., `geom_text()`)

```{r}
#| label: dvs-3-2

```

-   I can choose creative geometries (e.g., `geom_segment()`, `geom_ribbon)()`)

```{r}
#| label: dvs-3-3

```

**DVS-4: I can calculate numerical summaries of variables.**

-   Example using `summarize()` : Lab 3 Question 6

```{r}
#| label: dvs-4-summarize
teacher_evals_clean |>
  summarise(unique_instructors = n_distinct(teacher_id),
            unique_courses = n_distinct(course_id))
```

-   Example using `across()` :

```{r}
#| label: dvs-4-across

```

**DVS-5: I can find summaries of variables across multiple groups.**

-   Example 1: Lab 3 Question 11

```{r}
#| label: dvs-5-1
failing_courses <- teacher_evals_clean |>
  filter(seniority == 1) |>
  group_by(teacher_id) |>
  summarise(n_courses = n_distinct(course_id),
            avg_pct_fail = mean(percent_failed_cur, na.rm = TRUE),
            .groups = "drop")
```

-   Example 2: Lab 3 Question 9

```{r}
#| label: dvs-5-2
teacher_evals_clean |>
  group_by(course_id, teacher_id) |>
  summarise(n_questions = n_distinct(question_no), groups = "drop") |>
  filter(n_questions == 9) |>
  select(course_id, teacher_id)
```

**DVS-6: I can create tables which make my summaries clear to the reader.**

-   I can modify my column names to clearly communicate the data context: Challenge 3 Question 1

```{r}
#| label: dvs-6-ex-1
teacher_evals_compare <- teacher_evals |>
  filter(question_no == 903) |>
  mutate(SET_level = if_else(SET_score_avg >= 4, "excellent", "standard"),
         sen_level = case_when(seniority <= 4 ~ "junior",
                               seniority > 4 ~ "senior")) |>
  select(course_id, SET_level, sen_level)

teacher_evals_compare
```

-   I can modify the text in my table to be more readable (e.g., bold face for column headers): Challenge 2 Spicy

```{r}
#| label: dvs-6-ex-2
ggplot(data = surveys, mapping = aes(x = weight, y = species)) +
  geom_jitter(color = "steelblue", alpha = 0.1) +
  geom_boxplot(outlier.shape = NA) +
  labs(x = "Weight (g)", y = "Species", title = "Weight Distribution by Species") +
  theme(panel.grid.minor = element_blank()) +
  theme(axis.title.y = element_text(margin = margin(r = 10)))
```

-   I can arrange my table to have an intuitive ordering: Lab 4 Question 6

```{r}
#| label: dvs-6-ex-3
infant_cost_2018 <- ca_childcare |>
  filter(study_year == 2018) |>
  group_by(census_region) |>
  summarize(med_mc_infant = median(mc_infant, na.rm = TRUE),
            .groups = "drop") |>
  arrange(desc(med_mc_infant))

infant_cost_2018
```

**DVS-7: I show creativity in my tables.**

-   I can use non-default colors

```{r}
#| label: dvs-7-ex-1

```

-   I can modify the layout of my table to be more readable (e.g., `pivot_longer()` or `pivot_wider()`) : Challenge 4

```{r}
#| label: dvs-7-ex-2
med_childcare_cost_by_setting <- ca_childcare |>
  select(study_year, mcsa, mfccsa) |>
  pivot_longer(cols = c(mcsa, mfccsa),
               names_to = "setting",
               values_to = "cost") |>
  group_by(setting, study_year) |>
  summarise(median_cost = median(cost, na.rm = TRUE),
            .groups = "drop") |>
  pivot_wider(names_from = study_year,
              values_from = median_cost,
              names_prefix = "year_") |>
  arrange(setting, desc(year_2018))

med_childcare_cost_by_setting
```

## Program Efficiency

**PE-1: I can write concise code which does not repeat itself.**

-   using a single function call with multiple inputs (rather than multiple function calls): Lab 4 Question 5

```{r}
#| label: pe-1-one-call
med_income_by_region <- ca_childcare |>
  select(census_region, study_year, mhi_2018) |>
  group_by(census_region, study_year) |>
  summarize(median_income = median(mhi_2018, 
                                   na.rm = TRUE), 
            .groups = "drop") |>
  pivot_wider(names_from = study_year,
              values_from = median_income,
              names_prefix = "year_") |>
  select(census_region, year_2008, year_2018) |>
  arrange(desc(year_2018)) |>
  rename(median_income_2008 = year_2008,
         median_income_2018 = year_2018)

med_income_by_region
```

-   using `across()` :

```{r}
#| label: pe-1-across

```

-   using functions from the `map()` family

```{r}
#| label: pe-1-map-1

```

**PE-2: I can write functions to reduce repetition in my code.**

-   Example 1: Function that operates on vectors: Lab 3 Question 12

```{r}
#| label: pe-2-1
resp_avg <- teacher_evals_clean |>
  filter(academic_degree %in% c("dr", "prof")) |>
  group_by(teacher_id, seniority, sex) |>
  summarise(n_courses = n_distinct(course_id),
            avg_resp = mean(resp_share, na.rm = TRUE),
            .groups = "drop")

highest_resp_avg <- resp_avg |>
  filter(avg_resp == max(avg_resp))

lowest_resp_avg <- resp_avg |>
  filter(avg_resp == min(avg_resp))

highest_resp_avg
```

-   Example 2: Function that operates on data frames: Lab 4 Question 5

```{r}
#| label: pe-2-2
med_income_by_region <- ca_childcare |>
  select(census_region, study_year, mhi_2018) |>
  group_by(census_region, study_year) |>
  summarize(median_income = median(mhi_2018, 
                                   na.rm = TRUE), 
            .groups = "drop") |>
  pivot_wider(names_from = study_year,
              values_from = median_income,
              names_prefix = "year_") |>
  select(census_region, year_2008, year_2018) |>
  arrange(desc(year_2018)) |>
  rename(median_income_2008 = year_2008,
         median_income_2018 = year_2018)

med_income_by_region
```

-   Example 3: Function that operates on vectors *or* data frames: Challenge 4

```{r}
#| label: pe-2-3
med_childcare_cost_by_setting <- ca_childcare |>
  mutate(study_year = as.integer(study_year)) |>
  select(study_year, mcsa, mfccsa) |>
  pivot_longer(cols = c(mcsa, mfccsa),
               names_to = "setting",
               values_to = "cost") |>
  group_by(setting, study_year) |>
  summarise(median_cost = median(cost, na.rm = TRUE),
            .groups = "drop")

ggplot(med_childcare_cost_by_setting, aes(x = study_year,
                                          y = median_cost,
                                          color = setting)) +
    geom_line(size = 1) +
    geom_point(size = 2, alpha = 0.5) +
    scale_x_continuous(breaks = 
                         seq(min(med_childcare_cost_by_setting$study_year),
                        max(med_childcare_cost_by_setting$study_year), by = 1)) +
    labs(title = "Median Full-Time Weekly Childcare Costs in CA by Setting",
         x = "Study Year",
         y = "Median Weekly Cost ($)",
         color = "Setting") +
    theme_bw()
```

**PE-3:I can use iteration to reduce repetition in my code.**

-   using `across()`

```{r}
#| label: pe-3-across

```

-   using a `map()` function with **one** input (e.g., `map()`, `map_chr()`, `map_dbl()`, etc.)

```{r}
#| label: pe-3-map-1

```

-   using a `map()` function with **more than one** input (e.g., `map_2()` or `pmap()`)

```{r}
#| label: pe-3-map-2

```

**PE-4: I can use modern tools when carrying out my analysis.**

-   I can use functions which are not superseded or deprecated: Lab 3 Question 9

```{r}
#| label: pe-4-1
teacher_evals_clean |>
  group_by(course_id, teacher_id) |>
  summarise(n_questions = n_distinct(question_no), groups = "drop") |>
  filter(n_questions == 9) |>
  select(course_id, teacher_id)
```

-   I can connect a data wrangling pipeline into a `ggplot()` : Challenge 4

```{r}
#| label: pe-4-2
med_childcare_cost_by_setting <- ca_childcare |>
  mutate(study_year = as.integer(study_year)) |>
  select(study_year, mcsa, mfccsa) |>
  pivot_longer(cols = c(mcsa, mfccsa),
               names_to = "setting",
               values_to = "cost") |>
  group_by(setting, study_year) |>
  summarise(median_cost = median(cost, na.rm = TRUE),
            .groups = "drop")

ggplot(med_childcare_cost_by_setting, aes(x = study_year,
                                          y = median_cost,
                                          color = setting)) +
    geom_line(size = 1) +
    geom_point(size = 2, alpha = 0.5) +
    scale_x_continuous(breaks = 
                         seq(min(med_childcare_cost_by_setting$study_year),
                        max(med_childcare_cost_by_setting$study_year), by = 1)) +
    labs(title = "Median Full-Time Weekly Childcare Costs in CA by Setting",
         x = "Study Year",
         y = "Median Weekly Cost ($)",
         color = "Setting") +
    theme_bw()
```

## Data Simulation & Statistical Models

**DSSM-1: I can simulate data from a *variety* of probability models.**

-   Example 1: Lab 3 Question 2

```{r}
#| label: dsm-1-1
ggplot(teacher_evals_compare, aes(x = sen_level, fill = SET_level)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("excellent" = "purple", "standard" = "orange")) +
  labs(x = "Years of Experience",
       y = "",
       title = "Evaluation of Teachers' Use of Activities",
       fill = "Evaluation Rating")
```

-   Example 2: Lab 2 Question 16

```{r}
#| label: dsm-1-2
ggplot(data = surveys, mapping = aes(x = weight, y = species)) +
  geom_boxplot(outlier.shape = NA) +
  labs(x = "Weight (g)", y = "Species", title = "Distribution of Species by Weight") +
  geom_jitter(color = "steelblue", alpha = 0.3)
```

**DSSM-2: I can conduct common statistical analyses in R.**

-   Example 1: Lab 2 Question 17

```{r}
#| label: dsm-2-1
species_mod <- aov(weight ~ species, data = surveys)
summary(species_mod)
```

-   Example 2: Lab 3 Question 3

```{r}
#| label: dsm-2-2
chisq.test(table(teacher_evals_compare$sen_level, teacher_evals_compare$SET_level))
```

-   Example 3: Lab 4 Question 9

```{r}
#| label: dsm-2-3
reg_mod1 <- lm(mc_infant ~ mhi_2018, data = ca_childcare)
summary(reg_mod1)
```

## Revising My Thinking

<!-- How did you revise your thinking throughout the course? How did you revise your thinking on the code examples you have provided in your portfolio? -->

I revised my thinking in several ways throughout the course. I learned about programming efficiency, and how to write code that doesn't repeat the same function call multiple times. I also learned how to minimize the amount of objects saved in my code. Additionally, I learned how to reformat tables and graphs by altering text type, size, color, ordering, and placement. I learned how to use the here() package to import data, and I learned how to use Quarto formatting options to surpress messages from loading in packages, etc.

<!-- For the revisions included in your Portfolio, to help me understand the nature of your revisions, please denote somehow the feedback I provided you (e.g., boldface, italics, colored text) before your revisions. -->

## Extending My Thinking

<!-- How did you extended your thinking throughout the course? How did you extend your thinking on the code examples you have provided in your portfolio? -->

I extended my thinking in multiple ways throughout the course. As I learned how to alter the text in my tables and graphs, I used that knowledge to create good-looking data outputs that are readable to a viewer. I additionally used my knowledge of working with data sets to make my legends correspond to ordering on my graphs. I used plot modifications to make visualizations clearer to the reader, and I used colors to show creativity in my visualizations.

## Peer Support & Collaboration

<!-- Include an image or a description of feedback you gave that you are proud of (either in a peer review or in Discord). -->

This is feedback I am proud of giving for a peer review:

"Overall, the code written was very organized and efficient. I especially liked the colors in the second graph. For consistency, axis labels should use either (g) or (grams). Additionally, a lower alpha level may be useful so that the box plots are more visible."

<!-- Include a description of how you grew as a collaborator through the weekly pair programming activities.   -->

I grew as a collaborator through the weekly pair programming activities. It was helpful to bounce ideas back and forth with a partner as we tackled various problems, allowing us to see the concept from another person's point of view and see how they would tackle it in comparison to how I would typically start the problem. This allowed me to understand various concepts and functions more deeply, as I would discuss their use and application with my partner as we debated which operation or function would work best for what we are trying to do.
